---
title: "Lab 5"
author: "Alessandra Puig-Santana"
date: "10/26/2021"
output: 
  html_document: 
    theme: cerulean
    number_sections: yes
    toc: yes
    code_folding: hide #this will put an interactive code button that will let the user see the code chunk if they was to see it
---

```{r setup, include = TRUE, warning = FALSE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
```

# Confidence Intervals 

In this section, we'll learn how to use the 't.test()' function as a way to return the confidence interval using the t-distribution. 

## Example 1

Create a mock sample of observation values and find the confidence interval. Then, we'll learn how to report that.

```{r}
pika_mass <- c(82, 94, 110, 70, 86, 99, 102, 320, 5000)
# do i think this is representative sample of the population? if not, reevaluate and think what it represents. You have thought about how it is distributed, etc.

pika_mean <- mean(pika_mass) #finding mean

pika_sd <- sd(pika_mass) #finding standard distribution

pika_ci <- t.test(pika_mass) #this t.test function will give you the confidence interval
# this doesn't look like what we've seen, it comes out as a list. A flexible way to store information that has different classes. This is testing the null hypothesis on a single sample. Running on the console, you will see the outputs. 
# t-value is how different is your value from the null 0? 

pika_ci
```
### Risky and tedious way:
Mean pika mass at Paiute Pass is 91.9 g $\pm$ 13.5 g (mean $\pm$ 1 standard deviation, n = 7) with a 95% confidence interval of [79.4, 104.4] g. 

### A high initial investment, but safer and better way! 

In-line code referencing to stored objects! 

Use a single backtick on either side of a lowercase r to create an in-line bit of code.

Mean pika mass at Paiute Pass is `r round(pika_mean, 1)` g $\pm$ `r round(pika_sd, 1)` g (mean $\pm$ 1 standard deviation, n = `r length(pika_mass)`) with a 95% confidence interval of [`r pika_ci$conf.int[1]`, `r pika_ci$conf.int[2]`] g.

to call out data, go to console and use $. example: pika_ci$conf.int to get the confidence interval.

indexing: to pull individual values, use brackets. example: pika_mass[2] to pull the second value of the masses. 
