---
title: "Lab 5"
author: "Alessandra Puig-Santana"
date: "10/26/2021"
output: 
  html_document: 
    theme: cerulean
    number_sections: yes
    toc: yes
    code_folding: hide #this will put an interactive code button that will let the user see the code chunk if they was to see it
---

```{r setup, include = TRUE, warning = FALSE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
```

# Confidence Intervals 

In this section, we'll learn how to use the 't.test()' function as a way to return the confidence interval using the t-distribution. 

## Example 1

Create a mock sample of observation values and find the confidence interval. Then, we'll learn how to report that.

```{r}
pika_mass <- c(82, 94, 110, 70, 86, 99, 102, 320, 5000)
# do i think this is representative sample of the population? if not, reevaluate and think what it represents. You have thought about how it is distributed, etc.

pika_mean <- mean(pika_mass) #finding mean

pika_sd <- sd(pika_mass) #finding standard distribution

pika_ci <- t.test(pika_mass) #this t.test function will give you the confidence interval
# this doesn't look like what we've seen, it comes out as a list. A flexible way to store information that has different classes. This is testing the null hypothesis on a single sample. Running on the console, you will see the outputs. 
# t-value is how different is your value from the null 0? 

pika_ci
```
### Risky and tedious way:
Mean pika mass at Paiute Pass is 91.9 g $\pm$ 13.5 g (mean $\pm$ 1 standard deviation, n = 7) with a 95% confidence interval of [79.4, 104.4] g. 

### A high initial investment, but safer and better way! 

In-line code referencing to stored objects! 

Use a single backtick on either side of a lowercase r to create an in-line bit of code.

Mean pika mass at Paiute Pass is `r round(pika_mean, 1)` g $\pm$ `r round(pika_sd, 1)` g (mean $\pm$ 1 standard deviation, n = `r length(pika_mass)`) with a 95% confidence interval of [`r round(pika_ci$conf.int[1], 1)`, `r round(pika_ci$conf.int[2], 1)`] g.

to call out data, go to console and use $. example: pika_ci$conf.int to get the confidence interval.

indexing: to pull individual values, use brackets. example: pika_mass[2] to pull the second value of the masses. 

for significant figures, use round() function and then do ", the amount of sig figs you want" example: round(pika_mass, 2). 

# Two-sample t-test

Use a two-sample t-test to test the null hypothesis that samples were drawn from populations with the same mean (difference in means = 0).

- H0: Means difference = 0
- HA: Means difference is NOT = 0

This is a two-sided t-test because no directionality is implied. 

## Example 1
```{r}
oak_trees <- c(29, 19, 22, 30, 35, 16, 20, 7)
  
pine_trees <- c(48, 32, 41, 40, 56, 70, 30, 19, 20)
```

Is there a significance difference in mean heights for oak and pine trees?

```{r}
trees_t <- t.test(oak_trees, pine_trees)
```

The p-value of `r trees_t$p.value` means that there is `r (trees_t$p.value) * 100`% chance of finding sample means *at least as different as those I found* by random chance if they were drawn from populations with the same mean height. This is the tail ends on the distribution code. In the t-distribution, we are seeing how far away it is from zero but we are asking a two-sided question, so our p-value gives use that. 

## Example 2

Using the 'mpg' dataset to compare city gas mileage for SUVs and compact cars.
unique() function allows you to see the different classes in your data set

```{r}
# Create a subset called compact_suv that only contains observations where the class is suv OR compact
# filter allows you to select a column 
# class in c() is selecting suv and compact only
compact_suv <- mpg %>%
  filter(class %in% c("suv", "compact"))
  
```


```{r}
ggplot(data = compact_suv, aes(x = cty)) +
  geom_histogram(bins = 12) + # bins = # reduces the amount of bins
  facet_wrap(~class) # this divides your histogram by the different classes

ggplot(data = compact_suv, aes(sample = cty)) +
  geom_qq() + # q-q plot
  facet_wrap(~class)
```

Summary table 
```{r}
car_stats <- compact_suv %>%
  group_by(class) %>%
  summarize(mean_city = mean(cty),
            sd_city = round(sd(cty), 1),
            sample_size = n())

car_stats
```
Is there a significant difference between compact and suv cars?
```{r}
compact_sample <- compact_suv %>%
  filter(class == "compact") %>%
  pull(cty) #pulls vectors

suv_sample <- compact_suv %>%
  filter(class == "suv") %>%
  pull(cty)

cars_t <- t.test(compact_sample, suv_sample)
```

stats speak: reject the null hypothesis of equal mean city gas mileages (reject null)

Mean gas mileage for compact cars and SUVs differs significantly (t(`r cars_t$parameters`) = `r cars_t$statistic`, p < 0.001, $\alpha$ = 0.05). 
